---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/"
)
```

# economiccomplexity <img src="https://pachamaltese.github.io/economiccomplexity/hexicon.svg" width=150 align="right" alt="sticker"/>

[![R build status](https://github.com/pachamaltese/economiccomplexity/workflows/R-CMD-check/badge.svg)](https://github.com/pachamaltese/economiccomplexity/actions?workflow=R-CMD-check)
[![Codecov test coverage](https://codecov.io/gh/pachamaltese/economiccomplexity/branch/master/graph/badge.svg)](https://codecov.io/gh/pachamaltese/economiccomplexity?branch=master)
[![CRAN status](https://www.r-pkg.org/badges/version/economiccomplexity)](https://cran.r-project.org/package=economiccomplexity)
[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
[![Lifecycle: maturing](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing)

A consistent tool to explore bipartite relations. These methods are useful for different kind of relations such as The Product Space <doi:10.7551/mitpress/9647.001.0001> and The Human Disease Network <doi:10.1073/pnas.0701361104>.

The functions within this package correspond to code implementations of the equations described in Mariani, et al. (2015) <doi:10.1140/epjb/e2015-60298-7>, Hausmann, et al. (2014) <doi:10.7551/mitpress/9647.001.0001>, and Hausmann, et al. (2005) <doi:10.3386/w11905>.

## Installation

```{r, eval = FALSE}
# Install stable version from CRAN
install.packages("economiccomplexity")

# Install stable version from GitHub
devtools::install_github("pachamaltese/economiccomplexity")
```

## Community guidelines

If you want to contribute to the software, report issues or problems with the software, please fork the repo and send me a Pull Request or open an issue. I'm happy to receive ideas and I would do my best to coordinate efforts and improve this package without reinventing the wheel.

If you seek support or have questions you can start a thread on the issues section, or you can email me but I prefer open issues as probably more users have the same questions as you.

## Code of conduct

Please note that this project is released with a [Contributor Code of Conduct](https://pachamaltese.github.io/economiccomplexity/CODE_OF_CONDUCT.html).
By participating in this project you agree to abide by its terms.
