---
title: "How to use this package"
author: "Mauricio Vargas"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
bibliography: ../inst/REFERENCES.bib
vignette: >
  %\VignetteIndexEntry{How to use this package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, cache = FALSE, echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(eval = TRUE, message = FALSE, warning = FALSE)
```

# Basic trade example

## Package data

I'll use this demo dataset to illustrate the essential usage of the functions
within this package.

```{r}
library(economiccomplexity)

ec_trade_1962
```

## Revealed comparative Advantage (RCA)

You can obtain RCA with `ec_rca()`.

```{r}
rca <- ec_rca(
  data = ec_trade_1962,
  c = "country",
  p = "product",
  v = "value"
)

# 5x5 preview
rca[1:5,1:3]
```

And also you can obtain it in tabular version.

```{r}
rca_tbl <- ec_rca(
  data = ec_trade_1962,
  c = "country",
  p = "product",
  v = "value",
  tbl = T
)

rca_tbl
```

Another possibility, *not used to build networks from bipartite relations*, is
to obtain RCA as a matrix or tibble without discretization.

```{r}
rca_decimal <- ec_rca(
  data = ec_trade_1962,
  c = "country",
  p = "product",
  v = "value",
  discrete = F
)

# 5x3 preview
rca_decimal[1:5,1:3]
```

```{r}
rca_decimal_tbl <- ec_rca(
  data = ec_trade_1962,
  c = "country",
  p = "product",
  v = "value",
  tbl = T,
  discrete = F
)

rca_decimal_tbl
```

## Complexity Measures

You can compute both Economic Complexity Index (ECI) and Product Complexity
Index (PCI) by using `ec_complexity_measures()`. The calculations methods are
*reflections*, *eigenvalues* and *fitness* (default). See [@atlas2014] and
[@measuringcomplexity2015] for the methodological details.

```{r}
cm_reflections <- ec_complexity_measures(
  rca = rca,
  method = "reflections",
  tbl = T
)

cm_reflections$complexity_index_c
cm_reflections$complexity_index_p
```

```{r}
cm_eigenvalues <- ec_complexity_measures(
  rca = rca,
  method = "eigenvalues",
  tbl = T
)

cm_eigenvalues$complexity_index_c
cm_eigenvalues$complexity_index_p
```

```{r}
cm_fitness <- ec_complexity_measures(
  rca = rca,
  method = "fitness",
  tbl = T
)

cm_fitness$complexity_index_c
cm_fitness$complexity_index_p
```

## Proximity

Proximity matrices are used to create both country-country and product-product
networks. Using `proximity_matrices()` is straightforward.

```{r}
pro <- ec_proximity(
  rca = rca,
  d = cm_fitness$diversity,
  u = cm_fitness$ubiquity,
  tbl = T
)

pro$proximity_c
pro$proximity_p
```

## Networks

The `proximity_networks()` function is designed to use `igraph` for the internal
computations and also to pass proximity-based networks to `igraph`, `ggraph`
or export to Cytoscape by saving the output as csv/tsv.

To create some reduced networks I'll use a high proximity cutoff, so that the networks are limited to the spanning trees for clarity and fast rendering.

```{r}
net <- ec_networks(
  pc = pro$proximity_c,
  pp = pro$proximity_p,
  cutoff_c = 1,
  cutoff_p = 1,
  tbl = T
)

net$network_c
net$network_p
```

Just two basic examples with `ggraph`.

```{r, fig.width=20, fig.height=12}
library(igraph)
library(ggraph)
library(magrittr)

set.seed(200100)

net$network_c %>%
  graph_from_data_frame(directed = F) %>%
  ggraph(layout = "kk") +
  geom_edge_link(aes(edge_alpha = value, edge_width = value),
                 edge_colour = "#a8a8a8") +
  geom_node_point(color = "darkslategray4", size = 8) +
  geom_node_text(aes(label = name), vjust = 2.2) +
  ggtitle("The Country Space") +
  theme_void()
```

```{r, fig.width=20, fig.height=12}
net$network_p %>%
  graph_from_data_frame(directed = F) %>%
  ggraph(layout = "kk") +
  geom_edge_link(aes(edge_alpha = value, edge_width = value),
                 edge_colour = "#a8a8a8") +
  geom_node_point(color = "darkslategray4", size = 4) +
  geom_node_text(aes(label = name), vjust = 2.2) +
  ggtitle("The Product Space") +
  theme_void()
```

# References
