---
title: "How to use this package"
author: "Mauricio Vargas"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
bibliography: ../inst/references.bib
vignette: >
  %\VignetteIndexEntry{How to use this package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, cache = FALSE, echo = FALSE, message = FALSE, warning = FALSE, tidy = FALSE}
knitr::opts_chunk$set(eval = TRUE, message = FALSE, warning = FALSE)
```

# Basic trade example

## Package data

I'll use this demo dataset to illustrate the essential usage of the functions within this package:

```{r}
library(economiccomplexity)

world_trade_2017
```

## Revealed comparative Advantage (RCA)

You can obtain RCA with `revealed_comparative_advantage()`:
```{r}
rca <- revealed_comparative_advantage(
  d = world_trade_2017, 
  c = "reporter_iso",
  p = "product_code", 
  v = "export_value_usd"
)

rca[1:5,1:5]
```

And also you can obtain it in tabular version:
```{r}
rca_tbl <- revealed_comparative_advantage(
  d = world_trade_2017, 
  c = "reporter_iso",
  p = "product_code",
  v = "export_value_usd",
  tbl_output = T
)

rca_tbl
```

Another possibility, *not used to build networks from bipartite relations*, is to obtain RCA as a matrix or tibble without discretization:

```{r}
rca_decimal <- revealed_comparative_advantage(
  d = world_trade_2017,
  c = "reporter_iso",
  p = "product_code",
  v = "export_value_usd",
  tbl_output = F,
  discrete = F
)

rca_decimal[1:5,1:5]
```

```{r}
rca_decimal_tbl <- revealed_comparative_advantage(
  d = world_trade_2017,
  c = "reporter_iso",
  p = "product_code",
  v = "export_value_usd",
  tbl_output = T,
  discrete = T
)

rca_decimal_tbl
```

## Complexity Measures

You can compute both Economic Complexity Index (ECI) and Product Complexity Index (PCI) by using `complexity_measures()`. The calculations methods are *reflections*, *eigenvalues* and *fitness* (default). See [@atlas2014] and [@measuringcomplexity2015] for the methodological details.

```{r}
complexity_measures_reflections <- complexity_measures(
  world_rca_2017, 
  method = "reflections", 
  tbl_output = T
)

complexity_measures_reflections$economic_complexity_index
complexity_measures_reflections$product_complexity_index
```

```{r}
complexity_measures_eigenvalues <- complexity_measures(
  world_rca_2017,
  method = "eigenvalues",
  tbl_output = T
)

complexity_measures_eigenvalues$economic_complexity_index
complexity_measures_eigenvalues$product_complexity_index
```

```{r}
complexity_measures_fitness <- complexity_measures(
  world_rca_2017,
  method = "fitness", 
  tbl_output = T
)

complexity_measures_fitness$economic_complexity_index
complexity_measures_fitness$product_complexity_index
```

## Proximity

```{r}
# proximity <- proximity_matrices(d = world_rca_2017,
#   diversity = complexity_measures_2017$diversity,
#   ubiquity = complexity_measures_2017$ubiquity)
# 
# networks <- networks(proximity_matrices_2017$countries_proximity,
#   proximity_matrices_2017$products_proximity, tbl_output = T)
```
