---
title: "How to use this package"
author: "Mauricio Vargas"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
bibliography: ../inst/REFERENCES.bib
vignette: >
  %\VignetteIndexEntry{How to use this package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, cache = FALSE, echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(eval = TRUE, message = FALSE, warning = FALSE)
```

# Basic trade example

## Package data

I'll use this demo dataset to illustrate the essential usage of the functions
within this package.

```{r}
library(economiccomplexity)
ec_trade_1962
```

## Revealed comparative Advantage (RCA)

You can obtain RCA with `ec_rca()`.

```{r}
r <- rca(ec_trade_1962)
r
```

Another possibility is to obtain RCA as a matrix or tibble without discretization.

```{r}
r2 <- rca(ec_trade_1962, discrete = F)
r2
```

## Complexity Measures

You can compute both Economic Complexity Index (ECI) and Product Complexity
Index (PCI) by using `ec_complexity_measures()`. The calculations methods are
*fitness* (default), *reflections* and *eigenvalues*. See [@atlas2014] and
[@measuringcomplexity2015] for the methodological details.

```{r}
cf <- complexity(r, method = "fitness")
cf
```

## Proximity

Proximity matrices are used to create both country-country and product-product
networks. Using `proximity_matrices()` is straightforward.

```{r}
p <- proximity(r, cf$country_diversity, cf$product_ubiquity)
p
```

## Networks

The `proximity_networks()` function is designed to use `igraph` for the internal
computations and also to pass proximity-based networks to `igraph`, `ggraph`
or export to Cytoscape by saving the output as csv/tsv.

To create some reduced networks I'll use a high proximity cutoff, so that the networks are limited to the spanning trees for clarity and fast rendering.

```{r}
# net <- ec_networks(
#   pc = pro$proximity_c,
#   pp = pro$proximity_p,
#   cutoff_c = 1,
#   cutoff_p = 1,
#   tbl = T
# )
# 
# net$network_c
# net$network_p
```

Just two basic examples with `ggraph`.

```{r, fig.width=20, fig.height=12}
# library(igraph)
# library(ggraph)
# library(magrittr)
# 
# set.seed(200100)
# 
# net$network_c %>%
#   graph_from_data_frame(directed = F) %>%
#   ggraph(layout = "kk") +
#   geom_edge_link(aes(edge_alpha = value, edge_width = value),
#                  edge_colour = "#a8a8a8") +
#   geom_node_point(color = "darkslategray4", size = 8) +
#   geom_node_text(aes(label = name), vjust = 2.2) +
#   ggtitle("The Country Space") +
#   theme_void()
# ```
# 
# ```{r, fig.width=20, fig.height=12}
# net$network_p %>%
#   graph_from_data_frame(directed = F) %>%
#   ggraph(layout = "kk") +
#   geom_edge_link(aes(edge_alpha = value, edge_width = value),
#                  edge_colour = "#a8a8a8") +
#   geom_node_point(color = "darkslategray4", size = 4) +
#   geom_node_text(aes(label = name), vjust = 2.2) +
#   ggtitle("The Product Space") +
#   theme_void()
```

# References
