---
title: "How to use this package"
author: "Mauricio Vargas"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
bibliography: ../inst/REFERENCES.bib
vignette: >
  %\VignetteIndexEntry{How to use this package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, cache = FALSE, echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(eval = TRUE, message = FALSE, warning = FALSE)
```

# Galactic Federation

This example is inspired after the visualizations from @atlas2014 but heavily
inspired after Rick and Morty to keep it simple and lightweight.

## Package data

I'll use the Galactic Federation dataset to illustrate the essential usage of the functions
within this package.

```{r}
library(economiccomplexity)

galactic_federation
```

## Balassa Index

You can obtain Balassa Index with `balassa_index()`.

```{r}
bi <- balassa_index(
  data = galactic_federation,
  country = "planet",
  product = "product",
  value = "export_value"
)

bi
```

Another possibility is to obtain Balassa Index without discretization.

```{r}
bi_dec <- balassa_index(
  data =galactic_federation,
  country = "planet",
  product = "product",
  value = "export_value",
  discrete = F
)

bi_dec
```

## Complexity Measures

You can compute complexity indexes (e.g. such as the Economic Complexity Index and Product Complexity Index) by using `complexity_measures()`. The calculations methods are *fitness* (default), *reflections*, *eigenvalues*. See [@measuringcomplexity2015] for the methodological details.

The eigenvalues also calls the reflections methods in order to correct the index sign
in some special cases when the correlation between the output from both methods is negative.

##  Complexity-Fitness

```{r}
com_fit <- complexity_measures(balassa_index = bi)

com_fit$complexity_index_country
com_fit$complexity_index_product
```

## Reflections

```{r}
com_ref <- complexity_measures(
  balassa_index = bi,
  method = "reflections"
)

com_ref$complexity_index_country
com_ref$complexity_index_product
```

## Eigenvalues

```{r}
com_eig <- complexity_measures(
  balassa_index = bi,
  method = "eigenvalues"
)

com_eig$complexity_index_country
com_eig$complexity_index_product
```

## Proximity

Proximity matrices are used to create projections e.g. (country-country and product-product
networks) for bipartite networks. Using `proximity()` is straightforward.

```{r}
pro <- proximity(
  balassa_index = bi,
  balassa_sum_country = com_fit$balassa_sum_country,
  balassa_sum_product = com_fit$balassa_sum_product
)

pro$proximity_country
pro$proximity_product
```

## Networks

The `projections()` function is designed to use `igraph` for the internal
computations and also to pass proximity-based networks to `igraph`, `ggraph`
or export to Cytoscape by saving the output as csv/tsv.

```{r}
net <- projections(
  proximity_country = pro$proximity_country,
  proximity_product = pro$proximity_product,
  avg_links = 4,
  tolerance = 0.05
)

net$network_country
net$network_product
```

Just two basic examples with `ggraph`.
```{r, fig.width=7, fig.height=7}
set.seed(200100)
library(Matrix)
library(igraph)
library(ggraph)

aggregated_planets <- rowSums(galactic_federation)
V(net$network_country)$size <- aggregated_planets[match(V(net$network_country)$name, names(aggregated_planets))]

net$network_country %>%
  ggraph(layout = "kk") +
  geom_edge_link(aes(edge_width = weight), edge_colour = "#a8a8a8") +
  geom_node_point(aes(size = size), color = "darkslategray4") +
  geom_node_text(aes(label = name), vjust = 2.2) +
  ggtitle("Proximity Based Network Projection for Planets") +
  theme_void()
```

The network associated to

```{r, fig.width=7, fig.height=7}
set.seed(200100)

aggregated_products <- colSums(galactic_federation)
V(net$network_product)$size <- aggregated_products[match(V(net$network_product)$name, names(aggregated_products))]

net$network_product %>%
  ggraph(layout = "kk") +
  geom_edge_link(aes(edge_width = weight), edge_colour = "#a8a8a8") +
  geom_node_point(aes(size = size), color = "darkslategray4") +
  geom_node_text(aes(label = name), vjust = 2.2) +
  ggtitle("Proximity Based Network Projection for Products") +
  theme_void()
```

# References
