---
title: "How to use this package"
author: "Mauricio Vargas"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
bibliography: ../inst/REFERENCES.bib
vignette: >
  %\VignetteIndexEntry{How to use this package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, cache = FALSE, echo = FALSE, message = FALSE, warning = FALSE, tidy = FALSE}
knitr::opts_chunk$set(eval = TRUE, message = FALSE, warning = FALSE)
```

# Basic trade example

## Package data

I'll use this demo dataset to illustrate the essential usage of the functions within this package:

```{r}
library(economiccomplexity)

services_trade_2016$services_trade_2016_tibble
```

## Revealed comparative Advantage (RCA)

You can obtain RCA with `revealed_comparative_advantage()`:
```{r}
rca <- revealed_comparative_advantage(
  trade_data = services_trade_2016$services_trade_2016_tibble, 
  country = "country",
  product = "product", 
  value = "value"
)

# 5x5 preview
rca[1:5,1:3]
```

And also you can obtain it in tabular version:
```{r}
rca_tbl <- revealed_comparative_advantage(
  trade_data = services_trade_2016$services_trade_2016_tibble, 
  country = "country",
  product = "product", 
  value = "value",
  tbl_output = T
)

rca_tbl
```

Another possibility, *not used to build networks from bipartite relations*, is to obtain RCA as a matrix or tibble without discretization:

```{r}
rca_decimal <- revealed_comparative_advantage(
  trade_data = services_trade_2016$services_trade_2016_tibble, 
  country = "country",
  product = "product", 
  value = "value",
  discrete = F
)

# 5x3 preview
rca_decimal[1:5,1:3]
```

```{r}
rca_decimal_tbl <- revealed_comparative_advantage(
  trade_data = services_trade_2016$services_trade_2016_tibble, 
  country = "country",
  product = "product", 
  value = "value",
  tbl_output = T,
  discrete = F
)

rca_decimal_tbl
```

## Complexity Measures

You can compute both Economic Complexity Index (ECI) and Product Complexity Index (PCI) by using `complexity_measures()`. The calculations methods are *reflections*, *eigenvalues* and *fitness* (default). See [@atlas2014] and [@measuringcomplexity2015] for the methodological details.

```{r}
cm_reflections <- complexity_measures(
  rca, 
  method = "reflections", 
  tbl_output = T
)

cm_reflections$economic_complexity_index
cm_reflections$product_complexity_index
```

```{r}
cm_eigenvalues <- complexity_measures(
  rca,
  method = "eigenvalues",
  tbl_output = T
)

cm_eigenvalues$economic_complexity_index
cm_eigenvalues$product_complexity_index
```

```{r}
cm_fitness <- complexity_measures(
  rca,
  method = "fitness", 
  tbl_output = T
)

cm_fitness$economic_complexity_index
cm_fitness$product_complexity_index
```

## Proximity

Proximity matrices are used to create both country-country and product-product networks. Using `proximity_matrices()` is straightforward:
```{r}
pro <- proximity(
  rca,
  diversity = cm_fitness$diversity,
  ubiquity = cm_fitness$ubiquity,
  tbl_output = T
)

pro$proximity_countries
pro$proximity_products
```

## Networks

The `proximity_networks()` function is designed to use `igraph` for the internal computations and also to pass proximity-based networks to `igraph`, `ggraph` or export to Cytoscape by saving the output as csv/tsv.

To create some reduced networks I'll use a high proximity cutoff:

```{r}
net <- networks(
  pro$proximity_countries,
  pro$proximity_products,
  c_cutoff = 0.7,
  p_cutoff = 0.1,
  tbl_output = T
)

net$network_countries
net$network_products
```

Just a two examples with `ggraph`:

```{r, fig.width=20, fig.height=12}
library(igraph)
library(ggraph)
library(magrittr)

set.seed(200100)

net$network_countries %>% 
  graph_from_data_frame(directed = F) %>% 
  ggraph(layout = "kk") +
  geom_edge_link(aes(edge_alpha = value, edge_width = value), edge_colour = "#a8a8a8") +
  geom_node_point(color = "darkslategray4", size = 8) +
  geom_node_text(aes(label = name), vjust = 2.2) +
  ggtitle(
    expression(paste("Connections between",
                     italic(" countries "),
                     "based on the",
                     italic(" services "),
                     "they export"))) +
  theme_void()
```

```{r, fig.width=20, fig.height=12}
net$network_products %>% 
  graph_from_data_frame(directed = F) %>% 
  ggraph(layout = "kk") +
  geom_edge_link(aes(edge_alpha = value, edge_width = value), edge_colour = "#a8a8a8") +
  geom_node_point(color = "darkslategray4", size = 8) +
  geom_node_text(aes(label = name), vjust = 2.2) +
  ggtitle(
    expression(paste("Connections between",
                     italic(" services "),
                     "based on the",
                     italic(" countries "),
                     "that export them"))) +
  theme_void()
```

# References
